---
title: "polygonsplot"
author: 
  - name: Davide Corso
    affiliation:
    - University of Naples Federico II
    email: davidecrs92@gmail.com
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('polygonsplot')`"
vignette: >
  %\VignetteIndexEntry{polygonsplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL ## Related to https://stat.ethz.ch/pipermail/bioc-devel/2020-April/016656.html
)
```

# Introduction

Polygons Plot is designed to visualize multivariate ranges as opposite to single
data points and it is able to show 3 to 6 variables simultaneously showcase the 
min-max and average for each. The concept of _polygons plot_ was developed by 
[Donato Giovannelli](https://www.donatogiovannelli.com/).

# Installation

You can install `r BiocStyle::Biocpkg("polygonsplot")` from *Bioconductor* with the
following code:

```{r, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}
BiocManager::install("polygonsplot")
```


# Usage

```{r library}
library(polygonsplot)
```

To use the `polygonsplot` package, there are just two mandatory parameters: 
- the input `data.frame` 
- the shape (`integer`, 3=Triangle, 4=Square, 5=Pentagon, 6=Hexagon)


```{r load-data}
df <- read.csv(system.file(file.path("extdata", "example.csv"), package="polygonsplot"))
df
```
## Shape: 3 - Triangle

```{r fig.width=7, fig.height=7 }
plot_triangle <- polygonplot(df, shape = 3, fillcolor = "#e56b6f", linecolor = "#b56576")
plot_triangle
```

## Shape: 4 - Square

```{r fig.width=7, fig.height=7 }
plot_square <- polygonplot(df, shape = 4, fillcolor = "#57cc99", linecolor = "#38a3a5")
plot_square
```

## Shape: 5 - Pentagon

```{r fig.width=7, fig.height=7}
plot_pentagon <- polygonplot(df, shape = 5, fillcolor = "#c89f9c", linecolor = "#b36a5e")
plot_pentagon
```

## Shape: 6 - Hexagon

```{r fig.width=7, fig.height=7}
plot_hexagon <- polygonplot(df, shape = 6, fillcolor = "#0077b6", linecolor = "#023e8a")
plot_hexagon
```


## Input `data.frame` description

The input `data.frame` should be composed as follows:
- The first column should contain the strings `axis_min`, `axis_max` and `data` 
in that row order.
- All the columns, starting from the second, must be dedicated to the variables
of interest. In particular, the first two rows should contain the values for the 
`axis_min` and `axis_max` respectively. If left empty, the 
axis_min/max will be automatically calculated based on data given and 
extension will be decided based on the Axes range extension 
(`extra` parameter).
Starting from the third row, each variable column must contain all the data values.
- Column names will be used as labels of the axis 
(starting from the second one). Change them in order to change the labels, 
but don't leave them empty.

### Usage with a minimal example
```{r load-data-minimal}

df_min <- read.csv(system.file(file.path("extdata", "minimal_example.csv"), package="polygonsplot"))
df_min
```

```{r fig.width=7, fig.height=7}
plot_min <- polygonplot(df_min, shape = 6, fillcolor = "#f4d58d", linecolor = "#8d0801")
plot_min
```

The `polygonplot` function return a `ggplot2` object. 
`NOTE`: If you are going to change the theme of the returned ggplot object, it is
highly suggested to specify in the `theme` function the `aspect.ration = 1` in 
order to keep the text and the relative ticks aligned on the axis.


# `sessionInfo` {-}

<details><summary>Session info</summary>

```{r session_info, echo=FALSE, cache=FALSE}
Sys.time()
sessionInfo()
```

</details>