---
title: "polygonPlot"
author: 
  - name: Davide Corso
    affiliation:
    - University of Naples Federico II
    email: davidecrs92@gmail.com
  - name: Fang Huang
    affiliation:
    - CSIRO Mineral Resources, Australia
    email: fanghuang007@gmail.com
  - name: Anirudh Prabhu
    affiliation:
    - Carnegie Institution for Science
    email: aprabhu@carnegiescience.edu
  - name: Donato Giovannelli
    affiliation:
    - University of Naples Federico II
    email: donato.giovannelli@unina.it
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{polygonPlot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

----

# Introduction

_polygonPlot_ is an R package for visualizing multivariate ranges using 
polygon-based representations.
Unlike traditional scatter or radar plots that display individual data points, 
polygonPlot allows users to display min–max intervals and mean values across 3 to 6 variables simultaneously.

# Installation

You can install `polygonPlot` with the following code:

```{r, eval=FALSE}
devtools::install_github("giovannellilab/polygonPlot")
```


# Usage

## Libraries

```{r library}
library(polygonPlot)
library(ggplot2)
```

## Data Input Formats

There are two types of input accepted by the polygonPlot function.

1. The expected data.frame format of the function
2. Classic data.frame format.

Just be sure that the input object is `data.frame` type.

In both cases, the function returns a ggplot2 object, which includes the mean points of each axis connected by lines.

## 1. Expected data.frame format Structure
### Input Parameters
#### Required

The `polygonPlot` function requires two mandatory parameters: 

* __`df`__: the input `data.frame` that should be composed as follows:
  * The first column should contain the strings `axis_min`, `axis_max` and `data` 
in that row order.
  * All the columns, starting from the second, must be dedicated to the variables
of interest. In particular, the first two rows should contain the values for the 
`axis_min` and `axis_max` respectively. 
    * If left empty, the axis_min/max will be automatically calculated based on data given and 
extension will be decided based on the Axes range extension (`extra` parameter).
  * Starting from the third row, each variable column must contain all the data values.
  * As default, column names will be used as labels of the axis 
(starting from the second one). Change them in order to change the labels, 
but don't leave them empty.

* __`shape`__: the `integer` value indicating the shape of the polygon. 
Available numbers:
  * `3` for _Triangle_
  * `4` for _Square_
  * `5` for _Pentagon_
  * `6` for _Hexagon_

#### Optional

* __`extra`__:  axis range extension
* __`fillcolor`__ fill color of the polygon
* __`alpha`__ alpha value of the fill color
* __`linecolor`__ line color of the polygon border
* __`linetype`__ line type of the polygon border
* __`lwd`__ line width of the polygon border
* __`labels_axis`__ vector with the desired labels of the axis
* __`mean_line_color`__ color of the mean line inside the polygon
* __`annotation_tick_size`__ size of tick marks along the polygon axes
* __`annotation_label_size`__ size of the labels displayed along the polygon axes
* __`title`__ title of the plot
* __`fix_aspect_ratio`__ boolean flag to fix the aspect ratio of the plot as 
`1`. It is strongly recommended to leave it as default value `TRUE`. 
  * __NOTE__: If you are going to change the theme of the returned ggplot object, remember to put in 
the `theme` function the following code `aspect.ratio = 1` in order to 
keep the text and the relative ticks aligned on the axis. 

See the section `About the aspect.ratio` for a practical example.


Examples of an input `data.frame`
```{r load-data}
df <- read.csv(system.file(file.path("extdata", "example.csv"), package="polygonPlot"))
df
```

a minimal version
```{r load-data-minimal}

df <- read.csv(
  system.file(
    file.path(
      "extdata",
      "minimal_example.csv"
    ), 
    package="polygonPlot"
  )
)

df
```


## 2. Classic data.frame structure

In this case, you can use a regular data.frame as input and convert it using a helper function that prepares it for plotting.

```{r}
classic_df <- data.frame(
  data1=c(-300, 500, 300, 350),
  data2=c(4, 14, -2, 12),
  data3=c(0, 1000, 10, 50),
  data4=c(0, 50, 20, 40)
)

formatted_df <- polygonPlot::prepare_dataframe(classic_df)

# Through this function it becomes the first type of input
formatted_df
```




## Polygons
### Shape: 3 - Triangle

```{r fig.width=7, fig.height=7}
plot_triangle <- polygonPlot(df, shape = 3, fillcolor = "#e56b6f", linecolor = "#b56576",
                             labels_axis = c("earth", "moon", "sun"))
plot_triangle
```


### Shape: 4 - Square

```{r fig.width=7, fig.height=7}
plot_square <- polygonPlot(df, shape = 4, fillcolor = "#57cc99", linecolor = "#38a3a5", annotation_label_size = 8)
plot_square
```

### Shape: 5 - Pentagon

```{r fig.width=7, fig.height=7}
plot_pentagon <- polygonPlot(df, shape = 5, fillcolor = "#c89f9c", linecolor = "#b36a5e")
plot_pentagon
```

### Shape: 6 - Hexagon

```{r fig.width=7, fig.height=7}
plot_hexagon <- polygonPlot(df, shape = 6, fillcolor = "#0077b6", linecolor = "#023e8a", 
                            annotation_label_size = 8)
plot_hexagon
```


## About the `aspect.ratio`

Since `polygonPlot` returns a `ggplot2` object, the aspect ratio must be fixed to preserve the geometry of the polygon.  
When applying a custom theme, always include `theme(aspect.ratio = 1)` to maintain alignment between axes and labels.

```{r fig.width=7, fig.height=7}
plot_pentagon <- polygonPlot(df, shape = 5, fillcolor = "#c89f9c", linecolor = "#b36a5e")
plot_pentagon + theme_bw() + theme(aspect.ratio = 1)
```


### Overlay multiple polygons in one figure
To perform this task, the polygons must be pre-generated, each sharing the same minimum and maximum values of the axes.
This ensures that all polygons are drawn on a consistent scale, allowing for direct and meaningful visual comparison between datasets.

```{r}
# Definiamo i limiti comuni per ogni colonna
axis_min <- c(-300, -4, 0, 0, NA, -0.5)
axis_max <- c(500, 14, 1000, 50, NA, 0.5)

# Primo dataframe
df1 <- data.frame(
  info   = c("axis_min", "axis_max", "data", "data", "data"),
  data1  = c(axis_min[1], axis_max[1], 300, 460, 120),
  data2  = c(axis_min[2], axis_max[2], -2, 12, 4),
  data3  = c(axis_min[3], axis_max[3], 10, 270, 80),
  data4  = c(axis_min[4], axis_max[4], 20, 40, 15),
  data5  = c(axis_min[5], axis_max[5], 5, 10, 3),
  data6  = c(axis_min[6], axis_max[6], 0.1, 0.3, -0.2)
)

# Secondo dataframe (stessi axis_min e axis_max ma dati diversi)
df2 <- data.frame(
  info   = c("axis_min", "axis_max", "data", "data", "data"),
  data1  = c(axis_min[1], axis_max[1], 250, 380, 490),
  data2  = c(axis_min[2], axis_max[2], -3, 8, 10),
  data3  = c(axis_min[3], axis_max[3], 30, 200, 850),
  data4  = c(axis_min[4], axis_max[4], 10, 30, 45),
  data5  = c(axis_min[5], axis_max[5], 2, 6, 8),
  data6  = c(axis_min[6], axis_max[6], -0.1, 0.25, 0.4)
)

df1

df2



```


```{r}
p1 <- polygonPlot(df1, shape=4, fillcolor = "darkgreen", alpha=0.5, linecolor = "darkgreen", mean_line_color = "blue")

p1
```


```{r}
p2 <- polygonPlot(df2, shape=4, fillcolor = "orange", alpha=0.5, linecolor = "orange", mean_line_color = "red")

p2
```


Now that we have created the two plots, we can overlay them in a single chart.
To do this, we need to define:
- a list containing the plots
- the corresponding labels

The parameters for the `overlay_polygons` function are:

* __`plot_list`__ a list of polygon plots (ggplot2 objects). All plots should use the same axis ranges. The order of plots in this list determines which polygons appear in the foreground and which are in the background.
* __`label_list`__ is a vector of strings. Must have the same length of `plot_list` object
* __`legend_title`__ title of the legend
* __`include_mean_lines`__ logical flag indicating whether to include mean lines in the overlaid polygons

```{r}
plt_list <- list(p1,p2)
plt_labels <- c("Example1", "Example2")

overlay_polygons(plt_list, plt_labels, include_mean_lines = T)

```



Even with three plot we can see the differences. However, too many plots may provide a difficult visualization.
```{r}

# Limiti comuni (uguali ai precedenti)
axis_min <- c(-300, -4, 0, 0, NA, -0.5)
axis_max <- c(500, 14, 1000, 50, NA, 0.5)

# Terzo dataframe
df3 <- data.frame(
  info   = c("axis_min", "axis_max", "data", "data", "data"),
  data1  = c(axis_min[1], axis_max[1], 100, 250, 400),
  data2  = c(axis_min[2], axis_max[2], -1, 6, 10),
  data3  = c(axis_min[3], axis_max[3], 50, 300, 700),
  data4  = c(axis_min[4], axis_max[4], 5, 25, 35),
  data5  = c(axis_min[5], axis_max[5], 4, 8, 9),
  data6  = c(axis_min[6], axis_max[6], 0.0, 0.2, 0.45)
)

p3 <- polygonPlot(df3,   shape=4, fillcolor = "dodgerblue", alpha = 0.5, linecolor = "dodgerblue", mean_line_color = "yellow")
p3

```


```{r}

plt_list <- list(p1,p2,p3)
plt_labels <- c("Example1", "Example2", "Example3")

overlay_polygons(plt_list, plt_labels, include_mean_lines = T)

```




## Citation
Paper in preparation...


## `sessionInfo` {-}

<details><summary>Session info</summary>

```{r session_info, echo=FALSE, cache=FALSE}
Sys.time()
Sys.Date()
sessionInfo()
```

</details>
