---
title: "polygonPlot"
author: 
  - name: Davide Corso
    affiliation:
    - University of Naples Federico II
    email: davidecrs92@gmail.com
  - name: Fang Huang
    affiliation:
    - CSIRO Mineral Resources, Australia
    email: fanghuang007@gmail.com
  - name: Anirudh Prabhu
    affiliation:
    - Carnegie Institution for Science
    email: aprabhu@carnegiescience.edu
  - name: Donato Giovannelli
    affiliation:
    - University of Naples Federico II
    email: donato.giovannelli@unina.it
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{polygonPlot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

----

# Introduction

Polygon Plot is designed to visualize multivariate ranges as opposite to single
data points and it is able to show 3 to 6 variables simultaneously showcase the 
min-max and average for each. The concept of _polygon plot_ was developed by 
[Donato Giovannelli](https://www.donatogiovannelli.com/).

# Installation

You can install `polygonPlot` with the following code:
.
```{r, eval=FALSE}
devtools::install_github("giovannellilab/polygonPlot")
```


# Usage

## Libraries

```{r library}
library(polygonPlot)
library(ggplot2)
```


## Input Parameters
### Required

The `polygonplot` function requires two mandatory parameters: 

* __`df`__: the input `data.frame` that should be composed as follows:
  * The first column should contain the strings `axis_min`, `axis_max` and `data` 
in that row order.
  * All the columns, starting from the second, must be dedicated to the variables
of interest. In particular, the first two rows should contain the values for the 
`axis_min` and `axis_max` respectively. 
    * If left empty, the axis_min/max will be automatically calculated based on data given and 
extension will be decided based on the Axes range extension (`extra` parameter).
  * Starting from the third row, each variable column must contain all the data values.
  * As default, column names will be used as labels of the axis 
(starting from the second one). Change them in order to change the labels, 
but don't leave them empty.


* __`shape`__: the `integer` value indicating the shape of the polygon. 
Available numbers:
  * `3` for _Triangle_
  * `4` for _Square_
  * `5` for _Pentagon_
  * `6` for _Hexagon_


Examples of the input `data.frame`
```{r load-data}
df <- read.csv(system.file(file.path("extdata", "example.csv"), package="polygonPlot"))
df
```

a minimal version
```{r load-data-minimal}

df <- read.csv(
  system.file(
    file.path(
      "extdata",
      "minimal_example.csv"
    ), 
    package="polygonPlot"
  )
)

df
```

### Optional

* __`extra`__:  axis range extension
* __`fillcolor`__ fill color of the polygon
* __`alpha`__ alpha value of the fill color
* __`linecolor`__ line color of the polygon border
* __`linetype`__ line type of the polygon border
* __`lwd`__ line width of the polygon border
* __`labels_axis`__ vector with the desired labels of the axis
* __`title`__ title of the plot
* __`fix_aspect_ratio`__ boolean flag to fix the aspect ratio of the plot as 
`1`. It is strongly recommended to leave it as default value `TRUE`. 
  * __NOTE__: If you are going to change the theme of the returned ggplot object, remember to put in 
the `theme` function the following code `aspect.ratio = 1` in order to 
keep the text and the relative ticks aligned on the axis. 

See the section `About the aspect.ratio` for a practical example.

## Return

* __`ggplot2`__ object.


## Polygons
### Shape: 3 - Triangle

```{r fig.width=7, fig.height=7}
plot_triangle <- polygonplot(df, shape = 3, fillcolor = "#e56b6f", linecolor = "#b56576",
                             labels_axis = c("earth", "moon", "sun"))
plot_triangle
```


### Shape: 4 - Square

```{r fig.width=7, fig.height=7}
plot_square <- polygonplot(df, shape = 4, fillcolor = "#57cc99", linecolor = "#38a3a5", annotation_label_size = 8)
plot_square
```

### Shape: 5 - Pentagon

```{r fig.width=7, fig.height=7}
plot_pentagon <- polygonplot(df, shape = 5, fillcolor = "#c89f9c", linecolor = "#b36a5e")
plot_pentagon
```

### Shape: 6 - Hexagon

```{r fig.width=7, fig.height=7}
plot_hexagon <- polygonplot(df, shape = 6, fillcolor = "#0077b6", linecolor = "#023e8a", 
                            annotation_label_size = 8)
plot_hexagon
```


## About the `aspect.ratio`
As described above, the `polygonplot` function return a `ggplot2` object. 
If you are going to change the theme of the returned ggplot object, it is
strongly recommended to specify in the `theme` function the `aspect.ratio = 1` in 
order to keep the text and the relative ticks aligned on the axis.

```{r fig.width=7, fig.height=7}
plot_pentagon <- polygonplot(df, shape = 5, fillcolor = "#c89f9c", linecolor = "#b36a5e")
plot_pentagon + theme_bw() + theme(aspect.ratio = 1)
```


## `sessionInfo` {-}

<details><summary>Session info</summary>

```{r session_info, echo=FALSE, cache=FALSE}
Sys.time()
sessionInfo()
```

</details>
